Optimizing Structural Properties Of Neural Networks With Genetic Algorithms

### 2015-08-01 paper/paper.tex
add paper

@@

\paragraph{Parametric Random Networks -- Genotypes} Since we are not
interested in single network properties but rather want to find out which
structural elements of networks are important for learning success, we
introduced \emph{parametric random networks}, or \emph{network genotypes}.
Such a genotype $G$ consists of a set $\Theta$ of parameters that determine
which concrete networks are created when $G$ is expressed in a single
phenotype. Technically $G$ may be understood as a stochastic network
generator that returns a network $\mathcal{N}=G(\omega)$ for every seed
value $\omega$.
The genotypes' parameters $\theta\in\Theta$ were either integer or floating
point values and can be classified as affecting either (1) the topology of
$\omega\Ix{rec}$ or (2) the weight values of $\omega\Ix{rec}$ and
$\omega\Ix{fb}$. Important examples for (1) are the networks size $N$, the
occupation probability $p$ (for random Erdös-Renyi networks), the
neighborhood range $k$ (for ring topologies), or the number $l$ of layers
(for feed-forward networks). We furthermore introduced ratio parameters
that allowed us to interpolate between different topologies.  Parameters of
type (2) were the feedback strength \texttt{feedback} (scaling of
$\omega\Ix{fb}$) and the gain value \texttt{gain} (scaling of
$\omega\Ix{rec}$ by ).

@@

\paragraph{Tasks and Challenges}
For the optimization to work properly, the problems presented to the networks 
have (for the most part) to solvable by FORCE learning. While it was shown \citep{}
that large recurrent neural networks are capable of learning very complex functions
using the FORCE algorithm, the optimization requires a large number of learnings.
Therefore, the networks have to be restricted to being quite small ($\approx 100$ neurons)
and as such are less powerful.

To get a simple criterion to estimate whether it might be possible to learn a given function 
using FORCE with 100 neurons, we look at a single sinusoidal of frequency $\omega$ and 
determine the range of $\omega$ in which an (unoptimized) network (i.e. GAIN, 
FEEDBACK, p, Erdös Renyi) is able to learn the function (see fig. \ref{}).

\paragraph{Assessing the quality of reconstructed functions}
In order to optimize the function reconstruction of a force network, a measure $f(g, g^\prime)$ for
the concordance of target $g$ and network output $g^\prime$ is necessary. Simply using the difference
$f_d(g, g^\prime) = \left\| g-g^\prime \right\|_2$ has the drawback that small phase differences could 
produce huge differences in quality, even though $g^\prime$ has the exact same shape as $g$.

This measure becomes problematic for functions that are very close to zero for a longer time 
(order of chunk size), but since we are going to work with superpositions of sinusoidals of 
different frequencies, this drawback is unproblematic.

>>>>

    \newcommand{\param}[1]{\texttt{#1}}
    \newcommand{\feedback}{\param{feedback}}
    \newcommand{\gain}{\param{gain}}
    \newcommand{\size}{N}
    \newcommand{\ERprob}{p}
    \newcommand{\RINGk}{k}
    \newcommand{\FFlayers}{l}

@@

\paragraph{Tasks}
For the optimization to work properly, the problems presented to the networks 
have (for the most part) to solvable by FORCE learning. While it was shown \citep{}
that large recurrent neural networks are capable of learning very complex functions
using the FORCE algorithm, the optimization requires a large number of learnings.
Therefore, the networks have to be restricted to being quite small ($\approx 100$ neurons)
and as such are less powerful.

To get a simple criterion to estimate whether it might be possible to learn a given function 
using FORCE with 100 neurons, we look at a single sinusoidal of frequency $\omega$ and 
determine the range of $\omega$ in which an (unoptimized) network (i.e. \gain, 
\feedback, \ERprob) is able to learn the function (see fig. \ref{}).

\paragraph{Success}
In order to optimize the function reconstruction of a force network, a measure $f(g, g^\prime)$ for
the concordance of target $g$ and network output $g^\prime$ is necessary. Simply using the difference
$f_d(g, g^\prime) = \left\| g-g^\prime \right\|_2$ has the drawback that small phase differences could 
produce huge differences in quality, even though $g^\prime$ has the exact same shape as $g$.

This measure becomes problematic for functions that are very close to zero for a longer time 
(order of chunk size), but since we are going to work with superpositions of sinusoidals of 
different frequencies, this drawback is unproblematic.
\paragraph{Parametric Random Networks -- Genotypes} Since we are not
interested in single network properties but rather want to find out which
structural elements of networks are important for learning success, we
introduced \emph{parametric random networks}, or \emph{network genotypes}.
Such a genotype $G$ consists of a set $\Theta$ of parameters that determine
which concrete networks are created when $G$ is expressed in a single
phenotype. Technically $G$ may be understood as a stochastic network
generator that returns a network $\mathcal{N}=G(\omega)$ for every seed
value $\omega$.
The genotypes' parameters $\theta\in\Theta$ were either integer or floating
point values and can be classified as affecting either (1) the topology of
$\omega\Ix{rec}$ or (2) the weight values of $\omega\Ix{rec}$ and
$\omega\Ix{fb}$. Important examples for (1) are the networks size $N$, the
occupation probability $p$ (for random Erdös-Renyi networks), the
neighborhood range $k$ (for ring topologies), or the number $l$ of layers
(for feed-forward networks). We furthermore introduced ratio parameters
that allowed us to interpolate between different topologies.  Parameters of
type (2) were the feedback strength $\texttt{feedback}$ (scaling of
$\omega\Ix{fb}$) and the gain value \gain (scaling of
$\omega\Ix{rec}$ by ).

@@
